Figures

### 2016-01-01 scripts/figures.Rmd
Miseq ROC work done

sum.df<-read.csv("../data/process/2014-5-30/Variants/all.sum.csv",stringsAsFactors=F,comment.char = '#')

true_snv<-read.csv("../data/reference/PR8_WSN33.csv",comment.char = '#') # read in a csv of the differences in PR8 and WSN33 as determined by sanger
true_snv<-subset(true_snv,Ref!="-" & Allele.1!="-") # remove any indels as we are not concerned with them here
mutate(true_snv,mutant=paste0(Name,"_",Ref,Ref.Pos,Allele.1))->true_snv #make a column that has names the variants as segment_ReferencePositionVariant

PR8_var<-read.csv("../data/reference/PR8_variants.csv",comment.char = "#") # read in differences between plasmid control on viral PR8 used in dilutions

mutate(PR8_var,mutant=paste0(Name,"_",Ref,Ref.Pos,Allele.1))->PR8_var #make a column that has names the variants as segment_ReferencePositionVariant
sum.df<-mutate(sum.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants to variant calls
sum.df$category[sum.df$mutation %in% PR8_var$mutant]<-"PR8" # id the variants found in PR8
sum.df<-mutate(sum.df,exp.freq=find_freq.v(Id)) # add the expected frequency of the true variants

# Get information about the reference genome from the fasta file
reference.fasta<-"../data/reference/pHW2000_PR8-N_A.fa"
segments <- fasta.seqlengths(reference.fasta)
regions.bed <- data.frame(chr = gsub("[ ].*","", names(segments)), start=1, stop=segments, row.names=NULL)
regions.bed<-mutate(regions.bed,chr=as.character(chr))

possible_vars<-sum(regions.bed$stop)*3 # three possible variants/ position
id<-function(x){
  x<-strsplit(x,".",fixed=T)[[1]][1]
}

prior.seg.length<-c()

for(k in 1:length(regions.bed$chr)){ 
  prior.seg.length[k]<-sum(regions.bed$stop[1:k])  # the end positions of each segment relative to one sequence not including the trimming step
}

prior.seg.length<-c(0,prior.seg.length)

mutate(regions.bed,concat.start=start+prior.seg.length[match(chr,regions.bed$chr)],concat.stop=concat.start+stop)->regions.bed

melt(subset(regions.bed,select=-c(stop,start)),id.vars="chr")->regions.l

@@

  mutate(roc.df, exp.freq=find_freq.v(samp))-> roc.df
}  
tp<-unique(true_snv$mutant)
possible_tp<-length(tp)
adjust.coords<-function(roc.df,sum.df){ # adjsut the sensitivity and specificity called by pROC  
  samp<-roc.df$samp[1]
  samp.df<-subset(sum.df,Id==samp)
  sense.factor<-length(which(samp.df$category==T))/possible_tp
  TN.samp<-length(which(samp.df$category==F))
  mutate(roc.df,adj.sensitivity=sensitivity*sense.factor,FP=(TN.samp-TN.samp*specificity),TP=sense.factor*possible_tp,adj.specificity=(possible_vars-possible_tp-FP)/((possible_vars-possible_tp)))
}

@@

roc.df<-roc_df(roc.ls,"all")

roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df)
roc.df.adj<-roc.df.adj[order(roc.df.adj$adj.sensitivity),]
roc.df.adj<-mutate(roc.df.adj,exp.freq=as.factor(find_freq.v(samp)))
roc.df.adj$exp.freq <- factor(roc.df.adj$exp.freq, levels = rev(levels(roc.df.adj$exp.freq)))
miseq.roc<-ggplot(roc.df.adj,aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_brewer(palette = "Set1")+xlab("1-Specificity")+ylab("Sensitivity")+guides(color=guide_legend(title="Frequency"))+scale_y_continuous(limits=c(0,1))+scale_x_continuous(limits=c(0,0.001))

@@

@@

roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df)
roc.df.adj<-roc.df.adj[order(roc.df.adj$exp.freq,decreasing = T),]

knitr::kable(subset(roc.df.adj,select=c(exp.freq,TP,FP)))

>>>>

sum.df<-read.csv("../data/process/2014-5-30/one.sided/Variants/all.sum.csv",stringsAsFactors=F,comment.char = '#')

true_snv<-read.csv("../data/reference/PR8_WSN33.csv",comment.char = '#') # read in a csv of the differences in PR8 and WSN33 as determined by sanger
true_snv<-subset(true_snv,Ref!="-" & Allele.1!="-") # remove any indels as we are not concerned with them here
mutate(true_snv,mutant=paste0(Name,"_",Ref,Ref.Pos,Allele.1))->true_snv #make a column that has names the variants as segment_ReferencePositionVariant
true_snv<-rename(true_snv,c("Name"="chr","Ref.Pos"="pos")) # rename the columns for use later
PR8_var<-read.csv("../data/reference/PR8_variants.csv",comment.char = "#") # read in differences between plasmid control on viral PR8 used in dilutions

mutate(PR8_var,mutant=paste0(Name,"_",Ref,Ref.Pos,Allele.1))->PR8_var #make a column that has names the variants as segment_ReferencePositionVariant
PR8_var<-rename(PR8_var,c("Name"="chr","Ref.Pos"="pos")) # rename the columns for use later  
sum.df<-mutate(sum.df,category=mutation %in% true_snv$mutant) # add Column for True and false variants to variant calls
sum.df$category[sum.df$mutation %in% PR8_var$mutant]<-"PR8" # id the variants found in PR8
sum.df<-mutate(sum.df,exp.freq=find_freq.v(Id)) # add the expected frequency of the true variants

# Get information about the regions we can investigate based on the primer sites

regions.bed <-read.csv("../data/reference/pr8_noprimingsites.bed.csv",stringsAsFactors = F,comment.char = "#")
regions.bed<-mutate(regions.bed,length=stop-start-1) #because the numbers in the csv are the inner most binding sites of the primers and represent the last positiions that cannot be interogated.

possible_vars<-sum(regions.bed$length)*3 # three possible variants/ position

# limit the PR8 and the true positive data.frames to those not excluded by the priming sites

primer.cut<-function(x){ # a helper function to do this
  chr<-unique(x$chr)
  start<-regions.bed$start[match(x$chr,regions.bed$chr)]
  stop<-regions.bed$stop[match(x$chr,regions.bed$chr)]

  subset(x,pos>start & pos<stop)
}
sum.df<-ddply(sum.df,~chr,primer.cut) # interogating only the sites within primer location removes 244 called variants for two sided and 406 for one sided test
true_snv<-ddply(true_snv,~chr,primer.cut) # interogating only the sites within primer location removes 8 potential true positive (491)
PR8_var<-ddply(PR8_var,~chr,primer.cut) # interogating only the sites within primer location removes 12 potential PR8 differences (15)

id<-function(x){
  x<-strsplit(x,".",fixed=T)[[1]][1]
}

@@

  mutate(roc.df, exp.freq=find_freq.v(samp))-> roc.df
}  
tp<-unique(true_snv$mutant)
possible_tp<-length(tp)
adjust.coords<-function(roc.df,sum.df){ # adjust the sensitivity and specificity called by pROC  
  samp<-roc.df$samp[1]
  samp.df<-subset(sum.df,Id==samp)
  sense.factor<-length(which(samp.df$category==T))/possible_tp
  TN.samp<-length(which(samp.df$category==F))
  mutate(roc.df,adj.sensitivity=sensitivity*sense.factor,FP=(TN.samp-TN.samp*specificity),TP=sense.factor*possible_tp,adj.specificity=(possible_vars-possible_tp-FP)/((possible_vars-possible_tp)))
}

@@

roc.df<-roc_df(roc.ls)

roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df)
roc.df.adj<-roc.df.adj[order(roc.df.adj$adj.sensitivity),]
roc.df.adj<-mutate(roc.df.adj,exp.freq=as.factor(find_freq.v(samp)))
roc.df.adj$exp.freq <- factor(roc.df.adj$exp.freq, levels = rev(levels(roc.df.adj$exp.freq)))
miseq.roc<-ggplot(roc.df.adj,aes(x=1-adj.specificity,y=adj.sensitivity))+geom_step(aes(color=exp.freq),size=2)+scale_color_brewer(palette = "Set1")+xlab("1-Specificity")+ylab("Sensitivity")+guides(color=guide_legend(title="Frequency"))+scale_y_continuous(limits=c(0,1))#+scale_x_continuous(limits=c(0,0.001))

@@

pdf("../results/figures/miseq_roc.one.sided.pdf")
miseq.roc
dev.off()

@@

roc.df<-roc_df.one(roc.ls,0.001)
roc.df.adj<-ddply(roc.df,~samp,adjust.coords,sum.df)
roc.df.adj<-roc.df.adj[order(roc.df.adj$exp.freq,decreasing = T),]

knitr::kable(subset(roc.df.adj,select=c(exp.freq,adj.sensitivity,TP,adj.specificity,FP)))
